Hydra Battles and Cie

### 2021-09-07 doc/chapter-powers.tex
re-reading chapter 2 (partially)
first try
modifications of doc/movies/Makefile (includes additions and gaia)
Fir snippets in theories/additions

\begin{Coqsrc}
Section Definitions.

Variables (A: Type)
           (mult: A -> A -> A)
           (one: A).
Local Infix "*" := mult.
Local Notation "1" := one.
\end{Coqsrc}

\begin{Coqsrc}
Fixpoint power (x:A)(n:nat) : A :=
  match n with 
   | 0%nat => 1
   | S p =>   x * x ^ p
  end
where "x ^ n" := (power x n).

\begin{Coqsrc}
Compute power Z.mul 1%Z 2%Z 10.  
\end{Coqsrc}
\begin{Coqanswer}
   = 1024%Z
     : Z
\end{Coqanswer}

\begin{Coqsrc}
Open Scope string_scope.
Compute power append  "" "ab"  12.
\end{Coqsrc}

\begin{Coqanswer}
 = "abababababababababababab"
     : string
\end{Coqanswer}

@@

\begin{Coqsrc}
Compute N_bpow Z.mul 1%Z 2%Z 10. 
\end{Coqsrc}

\begin{Coqanswer}
 = 1024%Z
     : Z 
\end{Coqanswer}

\begin{Coqsrc}
Require Import String.
Open Scope string_scope.

Compute N_bpow append "" "ab"  12.
\end{Coqsrc}

\begin{Coqanswer}
  = "abababababababababababab"
     : string 
\end{Coqanswer}

@@

\begin{Coqsrc}
Module M2.
Section Definitions.

  Variables (A: Type) (zero one : A)  (plus mult  : A -> A -> A).

  Variable rt : semi_ring_theory  zero one plus mult   (@eq A).
  Add Ring Aring : rt.

  Notation "0" := zero.  
  Notation "1" := one.
  Notation "x + y" := (plus x y).  
  Notation "x * y " := (mult x y).

  Structure t : Type := mat{c00 : A;  c01 : A;  c10 : A;  c11 : A}.
\end{Coqsrc}

@@

\begin{Coqsrc}
Definition M2_mult (M M':t) : t := mat
  (c00 M * c00 M' + c01 M * c10 M') (c00 M * c01 M' + c01 M * c11 M')
  (c10 M * c00 M' + c11 M * c10 M') (c10 M * c01 M' + c11 M * c11 M').
\end{Coqsrc}

The neutral element for \texttt{M2\_mult} is the identity matrix.

\begin{Coqsrc}
Definition Id2 : t := mat 1 0 0 1.

End M2_Definitions.
End M2.
\end{Coqsrc}

@@

\begin{Coqsrc}
Fixpoint fib (n:nat) : N :=
  match n with
    0%nat | 1%nat => 1%N
  | S (S p as q) => fib p + fib q
  end.

Compute fib 20.
\end{Coqsrc}

\begin{Coqanswer}
= 10946 : N
\end{Coqanswer}

@@

\begin{exercise}
  \begin{enumerate}
  \item

@@

\begin{Coqsrc}
Definition mul2 (p q : N * N) :=
  match p, q with (a, b),(c,d) => (a*c + a*d + b*c, a*c + b*d) end.

Instance Mul2 : Monoid  mul2 (0,1).
(* Proof omitted *)
\end{Coqsrc}

The following lemma is a simplification of the equality~\vref{fibmat-eq1}.

\begin{Coqsrc}
Lemma next_fib (n:nat) : mul2 (1,0) (fib (S n), fib n) =
                         (fib (S (S n)), fib (S n)).  
\end{Coqsrc}

By induction\footnote{Simple induction, since $n$ has type \texttt{nat}.}
 over $n$, we obtain a the following equalities.

\begin{Coqsrc}
Definition fib_mul2 n := let (a,b) := power (M:=Mul2) (1,0) n   in (a+b).

Lemma fib_mul2_OK_0 (n:nat) : power (M:=Mul2) (1,0) (S (S n)) =
                              (fib (S n), fib n).

Lemma fib_mul2_OK n : fib n = fib_mul2 n.
\end{Coqsrc}

@@

\begin{Coqsrc}
Definition power_t := forall (A:Type)
                             (mult : A -> A -> A)
                             (one:A)
                             (x:A)
                             (n:N), A.
\end{Coqsrc}

@@

 Definition correct_expt_function (f : power_t) : Prop :=
  forall A (mult : A -> A -> A) (one:A)
            (x:A) (n:N), 
            power mult one x (N.to_nat n) = f A mult one x n.
\end{Coqbad}

@@

\begin{Coqbad}
Section CounterExample.
    Let mul (n p : nat) := n + 2 * p.
    Let one := 0.

    Remark mul_not_associative :
      exists  n p q,  mul n (mul p q) <> mul (mul n p) q.
    Proof. 
        exists 1, 1, 1; discriminate. 
    Qed.

    Remark one_not_neutral  :
      exists n : nat, mul one n <> n.
    Proof.
      exists 1; discriminate.
    Qed.

    Lemma correct_expt_too_strong : 
          ~ correct_expt_function (@N_bpow).
    Proof.
      intro H; specialize (H _ mul one 1  7%N).
      discriminate H.    
    Qed.

End CounterExample.
End Bad.
 \end{Coqbad}

@@

\begin{Coqsrc}
Class Mult_op (A:Type) := mult_op : A -> A -> A.  
\end{Coqsrc}

@@

@@

\begin{Coqsrc}
Delimit Scope M_scope with M.
Infix "*" := mult_op : M_scope.
Open Scope M_scope.  
\end{Coqsrc}

@@

\begin{Coqsrc}
Instance nat_mult_op : Mult_op nat  := Nat.mul.
\end{Coqsrc}

@@

\begin{Coqsrc}
Set Printing All.
Check  3 * 4. 
\end{Coqsrc}
\begin{Coqanswer}
@mult_op nat nat_mult_op (S (S (S O))) (S (S (S (S O))))
     : nat  
\end{Coqanswer}
\begin{Coqsrc}
Unset Printing All.
Compute 3 * 4.   
\end{Coqsrc}
\begin{Coqanswer}
 = 12  : nat  
\end{Coqanswer}

@@

\begin{Coqsrc}
Require Import String.

Instance string_op : Mult_op string := append.
Open Scope string_scope.

Example ex_string : "ab" * "cde" = "abcde".
Proof. reflexivity. Qed.
\end{Coqsrc}

@@

\begin{Coqsrc}
Class Monoid {A:Type}(op : Mult_op A)(one : A) : Prop :=
{
    op_assoc : forall x y z:A, x * (y * z) = x * y * z;
    one_left : forall x, one * x = x;
    one_right : forall x, x * one = x
}.
\end{Coqsrc}

@@

\begin{Coqsrc}
Instance Z_mult_op : Mult_op Z := Z.mul.

Instance ZMult : Monoid  Z_mult_op 1.
Proof. 
  split.
\end{Coqsrc}

\begin{Coqanswer}
3 subgoals, subgoal 1 (ID 8)

  ============================
   forall x y z : Z, (x * (y * z))%M = (x * y * z)%M

subgoal 2 (ID 9) is:
 forall x : Z, (1 * x)%M = x
subgoal 3 (ID 10) is:
 forall x : Z, (x * 1)%M = x}
\end{Coqanswer}

\begin{Coqsrc}
    all: unfold Z_mult_op, mult_op;intros;ring.
Qed.
\end{Coqsrc}

@@

The following function definition generates 9 sub-goals,
for proving that the measure on signatures is strictly decreasing along
the recursive calls. They are solved with the help of Standard Library's lemmas 
on arithmetic of \texttt{positive} numbers and Euclidean division.

>>>>

\inputsnippets{FirstSteps/Defs}

@@

\inputsnippets{FirstSteps/PowerCompute}

@@

\inputsnippets{FirstSteps/M2a}

@@

\inputsnippets{FirstSteps/M2Mult}

@@

\inputsnippets{Fib2/FibDef}

@@

\begin{exercise}
  \label{exercise:fibmat}
  \begin{enumerate}
  \item

@@

\inputsnippets{Fib2/mul2Def}
\inputsnippets{Fib2/mul2Monoid}

The following lemma is a simplification of the equality of Exercise~\ref{exercise:fibmat}.

\inputsnippets{Fib2/nextFib}

Let us consider a new definition of the Fibonacci function.

\inputsnippets{Fib2/fibMul2Def}
\inputsnippets{Fib2/fibMul2OK0}
\inputsnippets{Fib2/fibMul2OK}
\inputsnippets{Fib2/TimeFibMul2}

@@

\inputsnippets{FirstSteps/powerTDef}

@@

\inputsnippets{FirstSteps/Bada}

@@

\inputsnippets{FirstSteps/Badb}

@@

\inputsnippets{Monoid_def/MultOpClass}

@@

\inputsnippets{Monoid_def/MultOpEq}

@@

\inputsnippets{Monoid_def/MultOpInfix}

@@

\inputsnippets{Monoid_def/DemoNatMulta}

@@

\inputsnippets{Monoid_def/DemoNatMultb}

@@

\inputsnippets{Monoid_def/DemoStringMult}

@@

\inputsnippets{Monoid_def/MonoidClass}

@@

\inputsnippets{Monoid_instances/ZMultDef}

@@

The following function definition generates 9 subgoals,
for proving that the measure on signatures is strictly decreasing along
the recursive calls. They are solved with the help of Standard Library's lemmas 
on arithmetic of \texttt{positive} numbers and Euclidean division.
